### 2024-03-27 tesis_Ashish/Abstract/abstract.tex
update 27 Mar

The Pixel Cluster Counting (PCC) luminosity measurement method has emerged as a reliable offline technique for luminosity measurement developed for the CMS experiment.
%The PCC method is based on the rate of pixel clusters in the CMS pixel detector, which is the innermost part of the CMS tracker.
The PCC method is studied for its precision in luminosity measurement using 2017, 2018 and 2022  datasets. The results of these studies demonstrate that the PCC luminosity measurement method is stable and linear. %able to achieve final uncertainty on the luminosity measurement at the level of 1\% or better.
Final systematic uncertainty achieved on luminosity measurement for 2018 and 2022 datasets are 2.2\% and
is 1.5\% respectively. %1\% systematic uncertainty on the final luminosity measurement.
Moreover, the clusters and coincidences counting algorithms and their optimization for the Phase II High Luminosity LHC (HL-LHC) luminosity measurement are also developed, using simulated datasets for the proposed Phase II pixel detector. %This new pixel detector will replace the Run 2 pixel detector.
Linearity of the Tracker endcap pixel (TEPX) and Disk 4 Ring 1 luminometers are expected to be within 1\% (offline) up to a pileup of 200 and their statistical precision per second is about 0.1\% at pileup 200.
%It is expected to provide excellent statistical precision and linearity performance, which is essential for precision luminosity measurements at the HL-LHC.

>>>>

The Pixel Cluster Counting (PCC) luminosity measurement method has emerged as a reliable offline technique for luminosity measurement developed for the CMS experiment.
%The PCC method is based on the rate of pixel clusters in the CMS pixel detector, which is the innermost part of the CMS tracker.
The PCC method is studied for its precision in luminosity measurement using 2017, 2018 and 2022  datasets. The results of these studies demonstrate that the PCC luminosity measurement method is stable and linear. %able to achieve final uncertainty on the luminosity measurement at the level of 1\% or better.
Final systematic uncertainty achieved on luminosity measurement for 2018 and 2022 datasets are 2.2\% and
 1.4\% respectively. %1\% systematic uncertainty on the final luminosity measurement.
Moreover, the clusters and coincidences counting algorithms and their optimization for the Phase II High Luminosity LHC (HL-LHC) luminosity measurement are also developed, using simulated datasets for the proposed Phase II pixel detector. %This new pixel detector will replace the Run 2 pixel detector.
Linearity of the Tracker endcap pixel (TEPX) and Disk 4 Ring 1 luminometers is expected to be within 1\% (offline) up to a pileup of 200 and their statistical precision per second is about 0.1\% at pileup 200.
%It is expected to provide excellent statistical precision and linearity performance, which is essential for precision luminosity measurements at the HL-LHC.

### 2024-03-27 tesis_Ashish/Chapter1/chapter1.tex
update 27 Mar

Collider experiments are a type of particle physics experiment that involve that typically involve high energy collisions of electrons or protons. %colliding particles, typically protons or electrons, at very high energies.
These collisions can lead to the creation of new particles or the detection of new phenomena, providing insight into the fundamental nature of matter. Colliders are designed to accelerate particles to nearly the speed of light and guide them along a circular or linear path using magnetic fields. By colliding particles, physicists can study the resulting particle showers and detect the products of the collision using complex detectors. Two important parameters in the collider experiments are energy and luminosity. Collider experiments have played a crucial role in advancing our understanding of particle physics and continue to be a vital tool for scientific exploration.

@@

Luminosity is a measure of the number of particle collisions that occur per unit area per unit time. In particle physics, luminosity is a crucial parameter that determines the rate at which interesting particles may be produced in an collider experiment \cite{Herr:941318}.

The luminosity of a collider is given by the following equation:

\begin{equation}
L = \frac{N_1 N_2 f}{A} 
\end{equation}

where $N_1$ and $N_2$ are the number of particles in each beam, f is the frequency of bunch crossings, and A is a parameter that takes into account the overlap between the two beams.

The importance of luminosity in particle physics lies in its direct relationship with the production rate of particles of interest. The production rate of a given particle is the product of the luminosity and the cross-section of the process. Thus, luminosity directly affects the ability to detect rare events and make precise measurements.
%The more collisions that occur, the higher the probability of rare events happening, such as the creation of new particles or the detection of new phenomena.
Additionally, precise measurements require a large number of collisions to be able to statistically determine the properties of the particles being studied.

@@

\begin{equation}
R = L  \sigma,
\end{equation}

where %R is the production rate,
L is the luminosity, and $\sigma$ is the cross-section of the process that produces the particle. The cross-section is a measure of the probability that a given process will occur, and it is typically a very small number, on the order of picobarns ($10^{-36} cm^2$).
In order to achieve high luminosity, particle accelerators use various techniques to increase the number of particles in the beams and focus them to a small cross-sectional area. This is done by using superconducting magnets to guide the particles along a circular path and keep them confined to a small area. Another technique is to use radio frequency cavities to accelerate the particles to higher energies increasing the frequency of bunch crossings. In high-energy particle accelerators, such as the Large Hadron Collider (LHC) at CERN, the luminosity is typically measured in units of inverse femtobarns ($fb^{-1}$), which is a measure of the total number of collisions per unit area during the course of an experiment.
%For example, the LHC has a design luminosity of $10^{34} cm^{-2}s^{-1}$, which means that there are $10^{34}$ proton-proton collisions per square centimeter area per second.
The importance of luminosity can be seen in the many discoveries that have been made in particle physics due to high luminosity experiments. For example, the discovery of the Higgs boson at the LHC was made possible by the high luminosity of the collider, which allowed physicists to observe the rare Higgs boson signal over the background noise of other particles. In addition to discoveries, high luminosity experiments are also important for precision measurements. For example, the measurement of the properties of the top quark, the heaviest known elementary particle, requires a large number of collisions to be able to distinguish the signal from the background noise.

@@

In particle physics, the luminosity of colliding beams is a crucial measurement as it provides information about the rate at which rare processes occur and enables precise measurements of particle properties such as cross-section and couplings to fermions and bosons. Luminosity depends on several variables, including the number of protons in each beam, the size and shape of the beams, and their relative positioning. A slight change in any of these variables can significantly alter the luminosity.
%which in turn affects the interpretation of collision data.
Therefore, it is crucial to measure the luminosity with high precision.
%The measurement of luminosity with the pixel detector at the CMS experiment is particularly important  as it is a highly precise detector that is capable of measuring luminosity with great accuracy. It provides information about the number of charged particles passing through the detector, which is directly proportional to the number of collisions. This information is then used to calculate the luminosity. 
Improving the uncertainty in luminosity measurements is essential for many physics analyses.
%including searches for new particles and measurements of the properties of known particles.
It is a dominant factor in the measurement of Higgs Boson cross sections and couplings to other gauge bosons and fermions, as shown in Fig~\ref{fig:lum_unc}. As such, the uncertainty in cross-section and coupling measurements cannot be better than the uncertainty in luminosity measurement. Therefore, ongoing efforts are focused on improving the precision of luminosity measurements, with the pixel detector at CMS being a key component of this effort. By reducing the luminosity uncertainty, physicists can make more precise measurements of particle properties and potentially discover new physics beyond the Standard Model.

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.8\textwidth]{ashish_thesis/lumi_precision.png}
\caption[Phase II Uncertainty Projection]{ %
   Higgs boson production processes cross section projected uncertainty. In the most precisely measured Higgs boson production process, gluon fusion (ggH), luminosity uncertainty will dominate other experimental systematic uncertainty at the High Luminosity (HL)-LHC \cite{Dainese:2703572}.
}
\label{fig:lum_unc}
\end{figure}

@@

\section{Large Hadron Collider (LHC) and the CMS Experiment}

The CERN accelerator complex consists of a series of accelerators that work together to produce beams of particles at ever-increasing energies. These beams are then fed into the LHC for further acceleration and collision. The first step in the accelerator complex is the Linac 2, a linear accelerator that produces beams of protons with an energy of 50 MeV (mega-electronvolts). The protons are then sent to the Proton Synchrotron Booster (PSB), which accelerates them to an energy of 1.4 GeV (giga-electronvolts). From there, the proton beams are sent to the Proton Synchrotron (PS), which further accelerates them to an energy of 25 GeV. The next step in the complex is the Super Proton Synchrotron (SPS), which accelerates the protons to an energy of 450 GeV. The SPS also produces beams of heavy ions, such as lead nuclei, which are used for experiments in nuclear physics and other fields. Finally, the proton beams are sent to the LHC, where they are accelerated to their maximum energy of 6.5 TeV per proton.

LHC consists of a 27-kilometer ring of superconducting magnets, which guide two counter-rotating beams of protons to collide at four different locations around the ring, where four major experiments, ATLAS, the Compact Muon Solenoid (CMS) experiment, ALICE and LHCb are located as shown in Fig.~\ref{fig:lhc}. The protons are organized into "bunches," with each bunch containing around 100 billion protons. The bunches are spaced apart by 25 nanoseconds, and the LHC can contain up to 2,808 bunches per beam. The bunches in the LHC are designed to have a beam current of 0.55 amperes. It uses radiofrequency (RF) cavities to accelerate the proton beams. The RF frequency used is 400 MHz, which corresponds to a wavelength of 75 centimeters. The LHC uses quadrupole magnets to focus the proton beams and keep them tightly collimated. The LHC uses a total of 1,232 superconducting dipole magnets to bend the proton beams around the circular accelerator ring. The dipole magnets are cooled to a temperature of 1.9 Kelvin (-271.3°C) using liquid helium, and are capable of producing a magnetic field of up to 8.3 tesla. The LHC's detectors record data from proton-proton collisions at a rate of around 40 million bunch crossings per second, but only a tiny fraction of these are actually recorded for analysis. The LHC operates at a vacuum pressure of around $10^{-13}$ atmospheres, which is lower than the pressure on the surface of the moon.

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/lhc_schematic.png}
\caption[CERN Accelerator Complex]{%
    CERN accelerator complex consists of Linac 2, Proton Synchrotron Booster (PSB), Proton Synchrotron (PS), Super Proton Synchrotron (SPS) and Large Hadron Collider (LHC). LHC ring showing four interaction points where the ALICE, ATLAS, LHCb and CMS detectors are stationed \cite{lhc_complex}.
}
\label{fig:lhc}
\end{figure}

The CMS experiment is a multi-purpose particle detector that is designed to detect and study the particles produced in the collisions at the LHC. The CMS detector is a complex instrument that consists of several layers of sub-detectors, each designed to detect different types of particles and measure different properties of the particles. The CMS detector consists of a tracker, an electromagnetic calorimeter, hadron calorimeter and muon detector.

@@

\item The tracker is a cylindrical detector made up of two main subsystems: the pixel detector and the silicon strip detector. The pixel detector, the innermost subsystem near the interaction point, is crucial for precise track reconstruction and vertex identification. It includes four barrel layers and two endcap regions with hybrid pixel detector modules. These modules consist of a silicon sensor, segmented into roughly 100x150 $\mu$m² pixels, and readout electronics. When charged particles pass through the sensor, they create electron-hole pairs that induce a current, which is amplified, shaped, and digitized. The upgraded pixel detector used in Run2 contained over 120 million pixels, ensuring high granularity and hit efficiency. The pixel detector has a 4.8 $\mu$m  and 7.99 $\mu$m  spatial resolution along 100 $\mu$m and 150 $\mu$m pixel pitch respectively. The silicon strip detector surrounding the pixel detector provides extra tracking points for improved track reconstruction. It comprises of the Tracker Inner Barrel and Disks (TIB/TID) and the Tracker Outer Barrel and Endcaps (TOB/TEC). Silicon strip sensors, segmented into strips with a pitch of 80 to 205 $\mu$m, generate electron-hole pairs like the pixel detector. The signals are then processed by the readout electronics.
  %The silicon strip detector has a spatial resolution of 20-30 $\mu$m in the barrel region and 30-45 $\mu$m in the endcap region.

@@

\item The Electromagnetic Calorimeter (ECAL), a part of the CMS detector, is crucial for measuring the energy of photons and electrons from collisions. The ECAL contains over 75,000 lead tungstate (PbWO4) crystals arranged cylindrically around the interaction point. When photons or electrons hit the crystals, they generate a secondary particle shower, producing scintillation light proportional to the energy of the incoming particle. By assessing this light in each crystal, the ECAL can precisely determine the energy of the incoming particle. The ECAL covers a pseudorapidity range $|\eta| < 3.0$ \footnote{The pseudorapidity $\eta$ is calculated as $\eta = - \ln(\tan(\frac{\theta}{2}))$ and ranges from $-\infty$ to $+\infty$.}, facilitating measurement of particles at various angles to the beam axis. Its arrangement includes a barrel region and two endcap regions. The PbWO4 crystals vary in size and orientation in these regions to match particle trajectories. They are sufficiently long to absorb incoming particle energy. The ECAL boasts an energy resolution of less than 0.5\% for 50 GeV electrons and less than 1\% for 100 GeV photons. This precision results from the small crystal size, the PbWO4 crystals' high light yield, and advanced calibration procedures. Moreover, it has excellent timing resolution, crucial for differentiating between particles from the same collision and rejecting background events.

@@

\item The Hadron Calorimeter (HCAL) in the CMS detector, positioned outside of the Electromagnetic Calorimeter (ECAL), measures the energy of hadrons—particles comprised of quarks bound by the strong nuclear force. Since hadrons are abundantly produced in high-energy LHC collisions, the HCAL's design focuses on absorbing their energy for measurement. It is composed of alternating layers of brass, for energy absorption, and plastic scintillator, for light emission when particles pass through. Photodetectors convert this light into a measurable electrical signal. The Barrel HCAL, also known as HB in CMS, is the main part of the HCAL that is centered around the beam pipe and extends outwards. The pseudorapidity range for this section of the calorimeter is $|\eta|$ < 1.3. The Endcap HCAL, or HE, is designed to capture particles emitted at larger angles from the beam pipe, extending the coverage of the barrel region. The pseudorapidity range of the Endcap HCAL is 1.3 < $|\eta|$ < 3.0. The Forward HCAL, or HF, is situated at both ends of the detector to capture particles produced at the smallest angles to the beam direction. The pseudorapidity range of the Forward HCAL in the CMS experiment is 3.0 < $|\eta|$ < 5.2.

@@

\item The Muon Chambers are a vital part of the CMS detector. Designed to detect a variety of particles from high-energy collisions, the CMS is particularly equipped for detecting muons, highly penetrative charged particles that often evade other detection methods. Positioned on the outermost layer of the CMS detector, the muon chambers comprise three types of detectors: drift tubes (DTs), cathode strip chambers (CSCs), and resistive plate chambers (RPCs). The DTs are centrally located in the barrel region, while CSCs cover the forward and backward endcap regions. The RPCs, offering a quick trigger signal for muon detection, are found in both the barrel and endcap regions. These chambers detect muons by observing the ionization when a muon travels through the detector gas. By tracking muon trajectories, scientists can investigate the properties of these particles and explore physics beyond the Standard Model.

@@

During the operation of the CMS experiment, two main data-taking periods were conducted: Run 1, which spanned from 2010 to 2013, and Run 2, which took place from 2015 to 2018. During these periods, the CMS experiment measured the instantaneous and integrated luminosity of the proton-proton collisions produced by the LHC where the integrated luminosity for each year is shown in Fig.~\ref{fig:lumi}. During Run 1, the CMS experiment achieved a peak instantaneous luminosity of up to $7.6 \times 10^{33} cm^{-2} s^{-1}$, while during Run 2, the luminosity reached a peak value of up to $2.1 \times 10^{34} cm^{-2} s^{-1}$. The luminosity measurements during both runs were accompanied by careful assessments of the systematic uncertainties to ensure accurate and reliable results \cite{CMS-PAS-EWK-10-004, CMS-PAS-EWK-11-001, CMS-PAS-SMP-12-008, CMS-PAS-LUM-13-001, Sirunyan:2759951, pas_17, pas_18} as shown in Table. \ref{tab:lumi}.

@@

\begin{table}[h]
  \centering
  \caption[Integrated luminosity and its precision for Run1 and Run2]{Summary of center of mass energy, integrated luminosity, and systematic uncertainty per year during the CMS luminosity measurement.}
\begin{tabular}{c c c c}
  Year & Center of Mass Energy & Integrated Luminosity & Systematic Uncertainty \\
  & (TeV) & (fb$^{-1}$) & (\%) \\
\hline
2010 & 7 & 0.045 & 11 \\
2011 & 7 and 8 & 6.1 & 2.2 \\
2012 & 8 & 23.3 & 2.5 \\
2015 & 13 & 4.3 & 1.6 \\
2016 & 13 & 41.6 & 1.2 \\
2017 & 13 & 49.8 & 2.3 \\
2018 & 13 & 67.9 &  2.5\\
\end{tabular}
\label{tab:lumi}
\end{table}

@@

>>>>

Collider experiments are a type of particle physics experiment that typically involve high energy collisions of electrons or protons. %colliding particles, typically protons or electrons, at very high energies.
These collisions can lead to the creation of new particles or the detection of new phenomena, providing insight into the fundamental nature of matter. Colliders are designed to accelerate particles to nearly the speed of light and guide them along a circular or linear path using magnetic fields. By colliding particles, physicists can study the resulting particle showers and detect the products of the collision using complex detectors. Two important parameters in the collider experiments are energy and luminosity. Collider experiments have played a crucial role in advancing our understanding of particle physics and continue to be a vital tool for scientific exploration.

@@

Luminosity is a measure of the number of particle collisions that occur per unit area per unit time. In particle physics, luminosity is a crucial parameter that determines the rate at which interesting particles may be produced in a collider experiment \cite{Herr:941318}.

The instantaneous luminosity of a collider for head-on collisions is given by the following equation:

\begin{equation}
L_{inst} = \frac{N_1 N_2 f}{A} 
\end{equation}

where $N_1$ and $N_2$ are the number of particles in each beam, $f$ is the frequency of bunch crossings, and $A$ is a parameter that takes into account the overlap between the two beams that is the product of beam size.

The importance of luminosity in particle physics lies in its direct relationship with the production rate of particles of interest. The production rate of a given particle is the product of the luminosity and the cross-section of the process. Thus, luminosity directly affects the ability to detect rare events.
%The more collisions that occur, the higher the probability of rare events happening, such as the creation of new particles or the detection of new phenomena.
Additionally, precise measurements require a large number of collisions to be able to statistically determine the properties of the particles being studied.

@@

\begin{equation}
R = \sigma \: L_{inst} 
\end{equation}

where %R is the production rate,
$L_{inst}$ is the instantaneous luminosity, and $\sigma$ is the cross-section of the process that produces the particle. The cross-section is a measure of the probability that a given process will occur, and it is typically a very small number, on the order of picobarns ($10^{-36} cm^2$).
In order to achieve high luminosity, particle accelerators use various techniques to increase the number of particles in the beams and focus them to a small cross-sectional area. This is done by using superconducting magnets to guide the particles along a circular path and keep them confined to a small area. Another technique is to use radio frequency cavities to accelerate the particles to higher energies increasing the frequency of bunch crossings. In high-energy particle accelerators, such as the Large Hadron Collider (LHC) at CERN, the integrated luminosity is typically measured in units of inverse femtobarns ($fb^{-1}$), which is a measure of the total number of collisions per unit area during the course of an experiment.
%For example, the LHC has a design luminosity of $10^{34} cm^{-2}s^{-1}$, which means that there are $10^{34}$ proton-proton collisions per square centimeter area per second.
The importance of luminosity can be seen in the many discoveries that have been made in particle physics due to high luminosity experiments. For example, the discovery of the Higgs boson at the LHC was made possible by the high luminosity of the collider, which allowed physicists to observe the rare Higgs boson signal over the background noise of other particles. In addition to discoveries, high luminosity experiments are also important for precision measurements. For example, the precise measurement of the properties of the top quark, the heaviest known elementary particle, benefits significantly from the large number of collisions at the LHC. These collisions provide a robust dataset that reduces the uncertainty in the top quark mass measurement, allowing for more precise determination than were possible at the Tevatron.
%For example, the measurement of the properties of the top quark, the heaviest known elementary particle, requires a large number of collisions to be able to distinguish the signal from the background noise.

@@

In particle physics, the luminosity of colliding beams is a crucial measurement as it provides information about the rate at which rare processes occur and enables precise measurements of particle properties such as cross-section and couplings to fermions and bosons. Luminosity depends on several variables, including the number of protons in each beam, the size and shape of the beams, and their relative positioning. A slight change in any of these variables can significantly alter the luminosity.
%which in turn affects the interpretation of collision data.
Therefore, it is crucial to measure the luminosity with high precision.
%The measurement of luminosity with the pixel detector at the CMS experiment is particularly important  as it is a highly precise detector that is capable of measuring luminosity with great accuracy. It provides information about the number of charged particles passing through the detector, which is directly proportional to the number of collisions. This information is then used to calculate the luminosity. 
Improving the uncertainty in luminosity measurements is essential for many physics analyses.
%including searches for new particles and measurements of the properties of known particles.
It is a dominant factor in the measurement of Higgs Boson cross sections and couplings to other gauge bosons and fermions projected for High luminosity (HL)-LHC era, as shown in Fig~\ref{fig:lum_unc}. As such, the uncertainty in cross-section and coupling measurements cannot be better than the uncertainty in luminosity measurement. Therefore, ongoing efforts are focused on improving the precision of luminosity measurements, with the pixel detector at CMS being a key component of this effort. By reducing the luminosity uncertainty, physicists can make more precise measurements of particle properties and potentially discover new physics beyond the Standard Model.

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.8\textwidth]{ashish_thesis/lumi_precision.png}
\caption[Phase II Uncertainty Projection]{ %
   Higgs boson production processes cross section projected uncertainty. In the most precisely measured Higgs boson production process, gluon fusion (ggH), luminosity uncertainty will dominate other experimental systematic uncertainty at the HL-LHC \cite{Dainese:2703572}.
}
\label{fig:lum_unc}
\end{figure}

@@

\section{Large Hadron Collider and the CMS Experiment}

The CERN accelerator complex consists of a series of accelerators that work together to produce beams of particles at ever-increasing energies. These beams are then fed into the Large Hadron Collider (LHC) for further acceleration and collision. The first step in the accelerator complex is the Linac 2, a linear accelerator that produces beams of protons with an energy of 50 MeV (mega-electronvolts). The protons are then sent to the Proton Synchrotron Booster (PSB), which accelerates them to an energy of 1.4 GeV (giga-electronvolts). From there, the proton beams are sent to the Proton Synchrotron (PS), which further accelerates them to an energy of 25 GeV. The next step in the complex is the Super Proton Synchrotron (SPS), which accelerates the protons to an energy of 450 GeV. The SPS also produces beams of heavy ions, such as lead nuclei, which are used for experiments in nuclear physics and other fields. Finally, the proton beams are sent to the LHC, where they are accelerated to their maximum energy of 6.5 TeV per proton.

The LHC consists of a 27-kilometer ring of superconducting magnets, which guide two counter-rotating beams of protons to collide at four different locations around the ring, where four major experiments, ATLAS, the Compact Muon Solenoid (CMS) experiment, ALICE and LHCb are located as shown in Fig.~\ref{fig:lhc}. The protons are organized into "bunches," with each bunch containing around 100 billion protons. The bunches are spaced apart by 25 nanoseconds, and the LHC can contain up to 2,808 bunches per beam. The bunches in the LHC are designed to have a beam current of 0.55 amperes. It uses radiofrequency (RF) cavities to accelerate the proton beams. The RF frequency used is 400 MHz, which corresponds to a wavelength of 75 centimeters. The LHC uses quadrupole magnets to focus the proton beams and keep them tightly collimated. The LHC uses a total of 1,232 superconducting dipole magnets to bend the proton beams around the circular accelerator ring. The dipole magnets are cooled to a temperature of 1.9 Kelvin (-271.3°C) using liquid helium, and are capable of producing a magnetic field of up to 8.3 tesla. The LHC's detectors record data from proton-proton collisions at a rate of around 40 million bunch crossings per second, but only a tiny fraction of these are actually recorded for analysis. The LHC operates at a vacuum pressure of around $10^{-13}$ atmospheres, which is lower than the pressure on the surface of the moon.

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/lhc_schematic.png}
\caption[CERN Accelerator Complex]{%
    CERN accelerator complex consists of Linac 2, Proton Synchrotron Booster (PSB), Proton Synchrotron (PS), Super Proton Synchrotron (SPS) and Large Hadron Collider (LHC) which is the accelerator chain for proton physics at the LHC. LHC ring showing four interaction points where the ALICE, ATLAS, LHCb and CMS detectors are stationed \cite{lhc_complex}.
}
\label{fig:lhc}
\end{figure}

The CMS experiment \cite{CMS:2008xjf} is a multi-purpose particle detector that is designed to detect and study the particles produced in the collisions at the LHC. The CMS detector is a complex instrument that consists of several layers of sub-detectors, each designed to detect different types of particles and measure different properties of the particles. The CMS detector consists of a tracker, an electromagnetic calorimeter, hadron calorimeter, and a muon detector.

@@

\item The tracker is a cylindrical detector made up of two main subsystems: the pixel detector and the silicon strip detector \cite{CMS:2021ime}. The pixel detector, the innermost subsystem near the interaction point, is crucial for precise track reconstruction and vertex identification. It includes four barrel layers and two endcap regions with hybrid pixel detector modules. These modules consist of a silicon sensor, segmented into roughly 100x150 $\mu$m² pixels, and readout electronics. When charged particles pass through the sensor, they create electron-hole pairs that induce a current, which is amplified, shaped, and digitized. The upgraded pixel detector used in Run 2 contained over 120 million pixels, ensuring high granularity and hit efficiency. The pixel detector has a 4.8 $\mu$m  and 7.99 $\mu$m  spatial resolution along 100 $\mu$m and 150 $\mu$m pixel pitch respectively \cite{Dragicevic:2262932}. The silicon strip detector surrounding the pixel detector provides extra tracking points for improved track reconstruction. It comprises of the Tracker Inner Barrel and Disks (TIB/TID) and the Tracker Outer Barrel and Endcaps (TOB/TEC). Silicon strip sensors, segmented into strips with a pitch of 80 to 205 $\mu$m, generate electron-hole pairs like the pixel detector. The signals are then processed by the readout electronics.
  %The silicon strip detector has a spatial resolution of 20-30 $\mu$m in the barrel region and 30-45 $\mu$m in the endcap region.

@@

\item The Electromagnetic Calorimeter (ECAL), a part of the CMS detector, is crucial for measuring the energy of photons and electrons from collisions. The ECAL contains over 75,000 lead tungstate (PbWO4) crystals arranged cylindrically around the interaction point. When photons or electrons hit the crystals, they generate a secondary particle shower, producing scintillation light proportional to the energy of the incoming particle. By assessing this light in each crystal using avalanche photodiodes and vacuum phototriodes \cite{Hobson:2008zz}, the ECAL can precisely determine the energy of the incoming particle. The ECAL covers a pseudorapidity range $|\eta| < 3.0$ \footnote{The pseudorapidity $\eta$ is calculated as $\eta = - \ln(\tan(\frac{\theta}{2}))$ and ranges from $-\infty$ to $+\infty$.}, facilitating measurement of particles at various angles to the beam axis. Its arrangement includes a barrel region and two endcap regions. The PbWO4 crystals vary in size and orientation in these regions to match particle trajectories. They are sufficiently long to absorb incoming particle energy. The ECAL boasts an energy resolution of less than 0.5\% for 50 GeV electrons and less than 1\% for 100 GeV photons. This precision results from the small crystal size, the PbWO4 crystals' high light yield, and advanced calibration procedures. Moreover, it has excellent timing resolution, crucial for differentiating between particles from the same collision and rejecting background events.

@@

\item The Hadron Calorimeter (HCAL) \cite{Isik:2810162} in the CMS detector, positioned outside of the Electromagnetic Calorimeter (ECAL), measures the energy of hadrons—particles comprised of quarks bound by the strong nuclear force. Since hadrons are abundantly produced in high-energy LHC collisions, the HCAL design focuses on absorbing their energy for measurement. It is composed of alternating layers of brass, for energy absorption, and plastic scintillator, for light emission when particles pass through for the barrel and endcap calorimeters. HF uses quartz fibers and steel as active and passive components, respectively. Photodetectors convert this light into a measurable electrical signal. The Barrel HCAL, also known as HB in CMS, is the main part of the HCAL that is centered around the beam pipe and extends outwards. The pseudorapidity range for this section of the calorimeter is $|\eta|$ < 1.3. The Endcap HCAL, or HE, is designed to capture particles emitted at larger angles from the beam pipe, extending the coverage of the barrel region. The pseudorapidity range of the Endcap HCAL is 1.3 < $|\eta|$ < 3.0. The Forward HCAL, or HF, is situated at both ends of the detector to capture particles produced at the smallest angles to the beam direction. The pseudorapidity range of the Forward HCAL in the CMS experiment is 3.0 < $|\eta|$ < 5.2.

@@

\item The Muon Chambers \cite{CMS:2018rym} are a vital part of the CMS detector. Designed to detect a variety of particles from high-energy collisions, the CMS is particularly equipped for detecting muons, highly penetrative charged particles that often evade other detection methods. Positioned on the outermost layer of the CMS detector, the muon chambers comprise three types of detectors: drift tubes (DTs), cathode strip chambers (CSCs), and resistive plate chambers (RPCs). The DTs are centrally located in the barrel region, while CSCs cover the forward and backward endcap regions. The RPCs, offering a quick trigger signal for muon detection, are found in both the barrel and endcap regions. These chambers detect muons by observing the ionization when a muon travels through the detector gas. By tracking muon trajectories, scientists can investigate the properties of these particles and explore physics beyond the Standard Model.

@@

%During the operation of the CMS experiment, two main data-taking periods were conducted: Run 1, which spanned from 2010 to 2013, and Run 2, which took place from 2015 to 2018.
Since the beginning of the operation of the CMS experiment, three main data-taking periods have taken place: Run 1, Run 2, and Run 3, with the latter starting in 2022 and currently ongoing. During these periods, the CMS experiment measured the instantaneous and integrated luminosity of the proton-proton collisions produced by the LHC where the integrated luminosity for each year is shown in Fig.~\ref{fig:lumi}. During Run 1, the CMS experiment achieved a peak instantaneous luminosity of up to $7.6 \times 10^{33} cm^{-2} s^{-1}$, while during Run 2, the luminosity reached a peak value of up to $2.1 \times 10^{34} cm^{-2} s^{-1}$. The luminosity measurements during both runs were accompanied by careful assessments of the systematic uncertainties to ensure accurate and reliable results \cite{CMS-PAS-EWK-10-004, CMS-PAS-EWK-11-001, CMS-PAS-SMP-12-008, CMS-PAS-LUM-13-001, Sirunyan:2759951, pas_17, pas_18} as shown in Table. \ref{tab:lumi}. Luminosity systematic uncertaintities for heavy ion runs is shown in Table \ref{tab:lumi40}.

@@

\begin{table}[h]
  \centering
  \caption[Integrated luminosity and its precision for Run1 and Run2]{Summary of center of mass energy, integrated luminosity, and systematic uncertainty per year in CMS.}
\begin{tabular}{c c c c}
  Year & Center of Mass Energy & Integrated Luminosity & Systematic Uncertainty \\
  & (TeV) & (fb$^{-1}$) & (\%) \\
\hline
2010 & 7 & 0.045 & 11 \\
2011 & 7 and 8 & 6.1 & 2.2 \\
2012 & 8 & 23.3 & 2.5 \\
2015 & 13 & 4.3 & 1.6 \\
2016 & 13 & 41.6 & 1.2 \\
2017 & 13 & 49.8 & 2.3 \\
2018 & 13 & 67.9 &  2.5\\
\end{tabular}
\label{tab:lumi}
\end{table}

@@

\begin{table}[h]
  \centering
  \caption[Precision for Run1 and Run2]{Summary of center of mass energy, integrated luminosity, and systematic uncertainty per year during the CMS pPb and PbPb luminosity measurement \cite{CMS-PAS-LUM-13-002, CMS-PAS-LUM-17-002, CMS-PAS-LUM-18-001}.}
\begin{tabular}{c c c c}
  Year & Center of Mass Energy/nucleon & Systematic Uncertainty \\
  & (TeV) & (\%) \\
\hline
2013 & 2.76 & 3.4(3.6)  \\
2016 &  5.02  & 3.2(3.7)\\
2018 & 5.02 & 1.5 \\
\end{tabular}
\label{tab:lumi40}
\end{table}
